Tracing the Stellar Halo of Massive Galaxies to 100 kpc at $0.3<z<0.5$ using the Hyper Suprime-Cam

### 2017-05-25 draft/paper1_jenny.tex
Edit Introduction

    \update{
    Massive galaxies are important to our understanding of galaxy formation, and 
    their extended stellar halos contain crucial fossil records of their assembly 
    history. 
    In this work, we use ${\sim}100$ deg$^2$ of deep ($>28.5$ \sb{} in $i$-band), 
    high-quality (0.6\asec seeing) imaging data from the Hyper Suprime-Cam (HSC) 
    survey to investigate the surface mass density profiles and outer stellar halos 
    of massive galaxies. 
    Our sample consists of ${\sim}7000$ massive (\mstar{}$\geq 11.5$) central galaxies 
    at $0.3 < z < 0.5$. 
    The depth of the HSC survey reaches ${\sim}4$ magnitudes fainter than SDSS and 
    enables us to directly trace stellar mass distributions to 100 kpc without 
    requiring stacking. 
    Shallow images and/or over-simplified models like the \texttt{CModel} tend to 
    miss the extended stellar halo and result in underestimated stellar mass in a 
    mass-dependent way.  
    It will impact important applications like the measurements of the stellar mass 
    function. 
    With these deeper data, we explicitly show that at fixed stellar mass,
    massive central galaxies display surface mass density profiles that are not 
    self-similar, with the outer-slope becoming shallower and the fraction of mass 
    in the outskirt increasing with total stellar mass.
    We also show that the ellipticity of the stellar halo increases with stellar
    mass, while the optical color gradients do not strongly depend on stellar mass.
    Our results are broadly consistent with the two phase formation scenario of 
    massive galaxies in which outer envelopes are built up at late times from the 
    accretions of small systems. 
    The surface mass density profiles are provided in a convenient tabulated format 
    in order to facilitate comparisons with predictions from numerical simulations 
    of galaxy formation.
    }

@@

    Massive galaxies are unique laboratories for studying both cosmology and galaxy 
    formation. 
    Simulations of structure formation within the context of the $\Lambda$-CDM 
    cosmological model make predictions for the hierarchical growth of dark matter 
    halos and galaxies (e.g. \citealt{Baugh1996, DeLucia2006}), but there are many 
    open questions regarding the star-formation history, mass assembly process, and 
    structural evolution of massive galaxies. 
    Massive galaxies are thought to grow according to a ``two-phase'' formation 
    scenario (e.g. \citealt{Oser2010, Oser2012}). 
    In this scenario, the progenitors of $z{\sim} 0$ massive early-type galaxies 
    (ETGs) undergo a rapid growth phase at $z{\sim} 2$ triggered either by disk 
    instabilities or gas-rich mergers (e.g. \citealt{Hopkins2008, Dekel2009}). 
    Observationally, the progenitors of ETGs are thought to correspond to the 
    so-called ``red nuggets'' (e.g. \citealt{Damjanov2009}) which have smaller 
    effective radii ($R_{\mathrm{e}}$; e.g. \citealt{Trujillo2006, vanDokkum2008, 
    Cimatti2008}), slightly higher central velocity dispersion and stellar mass 
    density (\mden{}; e.g. \citealt{vandeSande2011, Belli2014}), and more disk-like 
    morphologies (e.g. \citealt{vanderWel2011}) than their local descendants 
    (e.g. \citealt{Bezanson2009, vanDokkum2010}).

    Following this initial epoch of early formation, feedback from stars and/or AGN 
    (e.g. \citealt{Sijacki2007, Fabian2012}) efficiently quench star formation in 
    massive galaxies and a large fraction of these massive progenitors become 
    quiescent at $z>1$ (e.g. \citealt{Bezanson2009, Kriek2016}). 
    The second phase in the ``two-phase'' formation scenario is driven by 
    non-dissipative processes such as dry mergers with other galaxies 
    (e.g. \citealt{Naab2006, Khochfar2006}). 
    \update{
    To explain the observed significant increase of $R_{\mathrm{e}}$ (e.g. 
    \citealt{Newman2012, vdWel2014}) as well as the build-up of the outer stellar 
    halos of ETGs (e.g. \citealt{Szomoru2012, Patel2013}), the ``two-phase'' 
    scenario suggests that minor mergers should dominate the mass assembly during 
    the second phase (e.g. \citealt{Hilz2012, Hilz2013, Oogi2013, Bedorf2013, 
    Laporte2013}).
    Both numerical simulation (e.g. \citealt{Oser2010}) and 
    semi-analytic modeling (SAM; e.g. \citealt{LeeYi2013, LeeYi2017}) agree that the 
    mass fraction in the accreted component should increase with total stellar mass, 
    although details still differ between models (e.g. \citealt{Lackner2012, Cooper2013, Qu2017}).
    According to recent results from the Illustris \footnote{
    \texttt{http://www.illustris-project.org/}} simulation 
    (\citealt{Vogelsberger2014}, \citealt{Genel2014}), the fraction of  
    accreted mass increases significantly with total stellar mass at \logms{}$>11.0$, 
    and the accreted component starts to dominate the stellar mass budget 
    ($f_{\mathrm{accreted}}>0.5$) at \logms{}$>11.5$ (\citealt{RodriguezGomez2016}).
    }
%% ------------------------------------------------------------------------------------ %%

    \update{
    Given the success of this two-phase scenario in explaining trends like overall size 
    growth, it is time to confront their predictions with additional observations using 
    local galaxy mass profiles. 
    Some progress has been made. 
    One-dimensional analysis has shown that a \ser{} law well-describes the surface 
    brightness (or stellar mass density) profiles of most nearby massive elliptical 
    galaxies (e.g. \citealt{Kormendy2009}; except for the most central region), with 
    the \ser{} index increasing with total luminosity (e.g. \citealt{Graham2013}).
	But, the situation may be more complicated, with a suggestion from 
	\citet{Schombert2015} that there are actually two families of light profiles. 
	Two-dimensional analysis, meanwhile, finds that the stellar distributions 
    of massive ETGs are better fit by multiple-component models 
    (e.g. \citealt{Huang2013a, Oh2017}). 
    To really confront the two-phase model requires very deep observations for large 
    samples of massive ETGs to correctly estimate their total stellar mass 
    (e.g. \citealt{Bernardi2013, DSouza2014}) and study their assembly history via 
    their accreted halos (e.g. \citealt{Capaccioli2015, Iodice2016, Iodice2017}).
    Currently, such work is limited by the depth of imaging from large-area surveys 
    such as the Sloan Digital Sky Survey (SDSS; e.g. \citealt{SDSSDR7, SDSSDR12}) or
    the small area of deeper imaging. 
    }

    In this paper, we take advantage of new high-resolution (median 
    FWHM${\sim} 0.6^{\arcsec}$ in $i$-band) and deep ($i$-band surface brightness 
    limit $> 28.5$\sb) imaging from the Hyper-Suprime Cam (HSC) Subaru Strategic 
    Program \citep[SSP,][]{HSCDR1} to characterize the outer parts of massive galaxies 
    as a function of their properties. 
    The deep imaging depth and excellent seeing conditions of HSC images make them 
    perfect for mapping the \mstar{} distributions of massive galaxies out to very 
    large radii. 
    \update{
    With the help of reliable spectroscopic and photometric redshift 
    (e.g. the ones for \redm{} clusters, \citealt{Rykoff2014}; \citealt{Rozo2015b}),
    we select a large sample (${\sim} 7000$) of massive central galaxies at 
    $0.3 < z < 0.5$ using ${\sim} 100$ square degree of data from the 
    HSC wide layer. 
    In this paper, we use this sample to 
    (a) measure individual \mden{} profiles for galaxies out to 100 kpc reliably with 
    the HSC data, 
    (b) examine the implications for the stellar mass function (SMF) when the outer 
    parts of galaxies are accounted for, and
    (c) investigate the dependence of their outer envelopes on stellar mass.    
    In a second paper in this series, we will further look into the potential 
    environmental dependence of structures in massive ETGs (Huang\etal in prep.).
    }

    %\jenny{I suggest removing this here. I do not think it belongs in the Intro:
    %Motivated by the ``two-phase'' scenario, and also due to the difficulty in 
    %providing appropriate model to describe the \mstar{} distribution of massive 
    %galaxy from its center to the very outskirt, we focus on \mstar{} within two 
    %different physical apertures in this work:}
    %\song{Since the original paper is quite long and complicated, we decide to define
    %the two aperture masses in the beginning, and explain why they are important.
    %For the new draft, I think it is Ok to do this later, but more comments on this 
    %are welcomed!}

%    \begin{itemize}
%    
%        \item The stellar mass within 10 kpc (hereafter noted \minn{}). 
%            According to the two-phase scenario, the ``in-situ'' star-formation quickly 
%            builds up the inner, dense core of massive ETGs.  
%            Based on recent observations and simulations (e.g.~\citealt{vanDokkum2010}; 
%            \citealt{RodriguezGomez2016}), the in-situ component of massive ETGs is 
%            mostly contained within one effective radius ($R_{\mathrm{e}}$, or 5-10 kpc). 
%            We use \minn{} as a proxy for the mass forged during the ``in-situ''
%            phase. 
%            Given the quality of the HSC data, we can reliably measure \minn{} over our 
%            redshift range ($1.0^{\arcsec}$ in radius equals 4.4 and 6.1 kpc at redshift
%            0.3 and 0.5).  
%            
%        \item The stellar mass within 100 kpc (hereafter noted \mtot{}). 
%            For our galaxy sample, a 100 kpc aperture corresponds to 5-10 
%            $\times R_{\mathrm{e}}$. 
%            We show in section \ref{ssec:mtotal} that most of the mass for these ETGs 
%            can be captured by a 100 kpc radius and that \mtot{} is a good proxy for the 
%            ``total'' \mstar{}. 
%            Although not perfect, we argue that our measurements of \mtot{} (which are 
%            actually measuring the light directly out to 100 kpc) are a better tracer 
%            of total \mstar{} than model-dependent results from shallower data 
%            (such as SDSS) which rely on extrapolating the light profiles of galaxies 
%            out to large radii.
%            
%       \end{itemize}}
%% ------------------------------------------------------------------------------------ %%

@@

    %$M_{\mathrm{halo}}$ denotes halo mass in general whereas $M_{\rm 200b}$ is 
    %explicitly defined as $M_{\rm 200b}\equiv M(<r_{\rm 200b})=200\bar{\rho} 
    %\frac{4}{3}\pi r_{\rm 200b}^3$ where $r_{\rm 200b}$
    %is the radius at which the mean interior density is equal to 200 times
    %the mean matter density ($\bar{\rho}$).

>>>>

    Massive galaxies are important to our understanding of galaxy formation, and 
    their extended stellar halos contain crucial fossil records of their assembly 
    history. 
    In this work, we use ${\sim}100$ deg$^2$ of deep ($>28.5$ \sb{} in $i$-band), 
    high-quality (0.6\asec seeing) imaging data from the Hyper Suprime-Cam (HSC) 
    survey to investigate the surface mass density profiles and outer stellar halos 
    of massive galaxies. 
    Our sample consists of ${\sim}7000$ massive (\mstar{}$\geq 11.5$) central galaxies 
    at $0.3 < z < 0.5$. 
    The depth of the HSC survey reaches ${\sim}4$ magnitudes fainter than SDSS and 
    enables us to directly trace stellar mass distributions to 100 kpc without 
    requiring stacking. 
    Shallow images and/or over-simplified models like the \texttt{CModel} tend to 
    miss the extended stellar halo and result in underestimated stellar mass in a 
    mass-dependent way.  
    It will impact important applications like the measurements of the stellar mass 
    function. 
    With these deeper data, we explicitly show that at fixed stellar mass,
    massive central galaxies display surface mass density profiles that are not 
    self-similar, with the outer-slope becoming shallower and the fraction of mass 
    in the outskirt increasing with total stellar mass.
    We also show that the ellipticity of the stellar halo increases with stellar
    mass, while the optical color gradients do not strongly depend on stellar mass.
    Our results are broadly consistent with the two phase formation scenario of 
    massive galaxies in which outer envelopes are built up at late times from the 
    accretions of small systems. 
    The surface mass density profiles are provided in a convenient tabulated format 
    in order to facilitate comparisons with predictions from numerical simulations 
    of galaxy formation.

@@

    Massive galaxies are fascinating objects that have experienced complex assembly 
    process and dramatic evolution.  
    They are crucial for both cosmology and the study of galaxy formation. 
    Simulations of structure formation within the context of the $\Lambda$-CDM 
    cosmological model make predictions for the hierarchical growth of dark matter 
    halos and galaxies (e.g. \citealt{Baugh1996, DeLucia2006}), but there are many 
    open questions regarding the star-formation history, mass assembly process, and 
    structural evolution of massive galaxies. 
    Massive galaxies are thought to grow according to a ``two-phase'' formation 
    scenario (e.g. \citealt{Oser2010, Oser2012}). 
    In this scenario, the progenitors of $z{\sim} 0$ massive early-type galaxies 
    (ETGs) undergo a rapid growth phase at $z{\sim} 2$ triggered either by disk 
    instabilities or gas-rich mergers (e.g. \citealt{Hopkins2008, Dekel2009}). 
    Observationally, the progenitors of ETGs are thought to correspond to the 
    population of massive quiescent galaxies at high redshift that have smaller 
    average effective radii ($R_{\mathrm{e}}$; e.g. \citealt{Trujillo2006, 
    vanDokkum2008, Cimatti2008}), slightly higher central velocity dispersion and 
    stellar mass density (\mden{}; e.g. \citealt{vandeSande2011, Belli2014}), and 
    more disk-like morphologies (e.g. \citealt{vanderWel2011}) than their local 
    descendants (e.g. \citealt{Bezanson2009, vanDokkum2010}) with similar stellar 
    mass.

    Following this initial epoch of early formation, feedback from stars and/or AGN 
    (e.g. \citealt{Sijacki2007, Fabian2012}) efficiently quench star formation in 
    massive galaxies. 
    A large fraction of these massive progenitors became quiescent at $z>1$ 
    (e.g. \citealt{Bezanson2009, Kriek2016}). 
    The second phase of their assembly is driven by non-dissipative processes such 
    as dry mergers with other galaxies (e.g. \citealt{Naab2006, Khochfar2006}). 
    To explain the observed increase of $R_{\mathrm{e}}$ (e.g. 
    \citealt{Newman2012, vdWel2014}) as well as the build-up of the outer stellar 
    halos of ETGs (e.g. \citealt{Szomoru2012, Patel2013}), the ``two-phase'' 
    scenario suggests that minor mergers should dominate the mass assembly during 
    the second phase (e.g. \citealt{Hilz2012, Hilz2013, Oogi2013, Bedorf2013, 
    Laporte2013}).
    Both numerical simulations (e.g. \citealt{Oser2010}) and 
    semi-analytic models (SAM; e.g. \citealt{LeeYi2013, LeeYi2017}) agree that the 
    mass fraction in the accreted component should increase with total stellar mass, 
    despite their technical differences (e.g. \citealt{Lackner2012, Cooper2013, 
    Qu2017}).
    For instance, recent result from the Illustris \footnote{
    \texttt{http://www.illustris-project.org/}} simulation 
    (\citealt{Vogelsberger2014}, \citealt{Genel2014}) predicts that the fraction of  
    accreted stars increases significantly with total stellar mass at \logms{}$>11.0$, 
    and the accreted component starts to dominate the stellar mass budget 
    ($f_{\mathrm{accreted}}>0.5$) at \logms{}$>11.5$ (\citealt{RodriguezGomez2016}).
%% ------------------------------------------------------------------------------------ %%

    Given the success of this two-phase scenario in explaining trends like overall size 
    growth, it is time to confront their predictions with additional observations, 
    especially the detailed profiles of surface mass density and shape for low redshift 
    massive galaxies.  
    Early one-dimensional analysis of nearby massive ETGs suggest that the 
    single-\ser{} profile can describe their surface brightness (or surface mass 
    density) profiles reasonably well (e.g. \citealt{Kormendy2009}; except for the
    most central region), with the \ser{} index increasing with total luminosity 
    (e.g. \citealt{Graham2013}).
	However, the situation is in fact more complicated, with a suggestion from 
	\citet{Schombert2015} that there are actually two families of light profiles:
	the ones that follow single-\ser{} law well and the ones show significant 
	deviation.
	Two-dimensional analysis, meanwhile, finds that the stellar distributions 
    of massive ETGs are often better fit by multiple-component models 
    (e.g. \citealt{Huang2013a, Oh2017}). 
    \citet{Huang2013b} further suggests a connection between the multi-component 
    nature of massive galaxies and their two-phase assembly history. 
    To further confront the two-phase scenario requires very deep observations of 
    large samples of massive ETGs to correctly estimate their total stellar mass 
    (e.g. \citealt{Bernardi2013, DSouza2014}) and explore their assembly history via 
    their extended, accreted halos (e.g. \citealt{Capaccioli2015, Iodice2016, 
    Iodice2017}).
    Currently, such work is limited by the depth of imaging from large-area surveys 
    such as the Sloan Digital Sky Survey (SDSS; e.g. \citealt{SDSSDR7, SDSSDR12}) or
    the small sample size of massive galaxies with deep images.

    In this paper, we take advantage of new high-quality (median 
    FWHM${\sim} 0.6^{\arcsec}$ in $i$-band) and deep ($i$-band surface brightness 
    limit $> 28.5$\sb) images from the Hyper Suprime-Cam (HSC) Subaru Strategic 
    Program \citep[SSP,][]{HSCDR1} to characterize the outer parts of massive galaxies 
    as a function of their properties. 
    The deep imaging depth and excellent seeing conditions of HSC images make them 
    perfect for mapping the \mstar{} distributions of massive galaxies out to very 
    large radii. 
    With the help of reliable spectroscopic and photometric redshift 
    (e.g. the ones for \redm{} clusters, \citealt{Rykoff2014}; \citealt{Rozo2015b}),
    we select a large sample (${\sim} 7000$) of massive central galaxies at 
    $0.3 < z < 0.5$ using ${\sim} 100$ deg$^2$ of data from the HSC wide layer. 
    In this paper, we use this sample to 
    (a) reliably measure individual surface mass density (\mden{}) profile of 
    massive galaxy out to 100 kpc, 
    (b) investigate the dependence of their outer stellar halos on total stellar 
    mass, and 
    (c) examine the implications for the stellar mass function (SMF) when the outer 
    halos of massive galaxies are accounted for.   
    In a second paper in this series, we will further look into the potential 
    environmental dependence of structures in massive ETGs (Huang\etal in prep.).

@@

    $M_{\mathrm{halo}}$ denotes halo mass in general whereas $M_{\rm 200b}$ is 
    explicitly defined as $M_{\rm 200b}\equiv M(<r_{\rm 200b})=200\bar{\rho} 
    \frac{4}{3}\pi r_{\rm 200b}^3$ where $r_{\rm 200b}$
    is the radius at which the mean interior density is equal to 200 times
    the mean matter density ($\bar{\rho}$).
